---
title: "Dppa2_4_public_data_integration"
author: "Zhiyuan Chen"
date: "`r format(Sys.time(), '%B %e, %Y')`"
output: 
  html_notebook:
    theme: paper 
    toc: yes
---


#### Meta data info

Reik_lab: http://genesdev.cshlp.org/content/33/3-4/194

  + Reverse_stranded PolyA RNA-seq
  + SE 50bp
  + WT x 2, Dppa2KO x 2, Dppa4KO x 2, Dppa2_4DKO x 1
  
Trono_lab: https://www.embopress.org/doi/full/10.15252/embr.201847382

  + Reverse_stranded total RNA-seq
  + PE 75bp
  + WT x 3, Dppa2KO x 3, Dppa4KO x 3
  
Wang_lab: https://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.3000324

  + PolyA RNA-seq (non-stranded)
  + PE 150bp
  + NC x 2, Dppa2KD x 2, Dppa4KD x 2, Dppa2_4DKD x 2, OE_CTRR x 2, Dppa2OE x 2
  
#### Import functions
```{r functions}
source("../script/utils.R")
```

#### Import RPKM for each datasets
##### Reik_lab RPKM
```{r input Reik RPKM}
Reik_RPKM_files <- list.files("../input/public/", 
                         pattern = "CRISPR*")

Reik_RPKM <- inputStringTieRPKMfiles(Reik_RPKM_files,
                                RPKMDataPath = "../input/public/")
#remove ERCC annotations
Reik_RPKM <- Reik_RPKM[-grep("ERCC", Reik_RPKM$id),]

#use simplified sample names 
tmp_Name <- c("reik_p2_ko1", "reik_p2_ko2", 
              "reik_p4_ko1", "reik_p4_ko2",
              "reik_dko", "reik_wt1", "reik_wt2")

colnames(Reik_RPKM) <- c("id", "name", tmp_Name)

#reorder columns
Reik_RPKM <- Reik_RPKM[, c("id", "name",
                           "reik_wt1", "reik_wt2", "reik_p2_ko1",
                           "reik_p2_ko2", "reik_p4_ko1", "reik_p4_ko2", 
                           "reik_dko")]
```

##### Trono_lab RPKM
```{r input Trono RPKM}
Trono_RPKM_files <- list.files("../input/public/", 
                         pattern = "RNAseq_*")

Trono_RPKM <- inputStringTieRPKMfiles(Trono_RPKM_files,
                                RPKMDataPath = "../input/public/")
#remove ERCC annotations
Trono_RPKM <- Trono_RPKM[-grep("ERCC", Trono_RPKM$id),]

#use simplified sample names 
tmp_Name <- c("trono_p2_ko1", "trono_p2_ko2", "trono_p2_ko3", 
              "trono_p4_ko1", "trono_p4_ko2", "trono_p4_ko3",
              "trono_wt1", "trono_wt2", "trono_wt3")

colnames(Trono_RPKM) <- c("id", "name", tmp_Name)

#reorder columns
Trono_RPKM <- Trono_RPKM[, c("id", "name",
                           "trono_wt1", "trono_wt2", "trono_wt3",
                           "trono_p2_ko1", "trono_p2_ko2", "trono_p2_ko3",
                           "trono_p4_ko1", "trono_p4_ko2", "trono_p4_ko3")]
```

##### Wang_lab RPKM
```{r input Wang RPKM}
Wang_RPKM_files1 <- list.files("../input/public/", 
                         pattern = "si_*")
Wang_RPKM_files2 <- list.files("../input/public/", 
                         pattern = "OE_*")
Wang_RPKM_files <- c(Wang_RPKM_files1, Wang_RPKM_files2)

Wang_RPKM <- inputStringTieRPKMfiles(Wang_RPKM_files,
                                RPKMDataPath = "../input/public/")
#remove ERCC annotations
Wang_RPKM <- Wang_RPKM[-grep("ERCC", Wang_RPKM$id),]
```

```{r input Wang RPKM2}
#use simplified sample names 
tmp_Name <- c("wang_p24_dkd1", "wang_p24_dkd2", 
              "wang_p2_kd1", "wang_p2_kd2", 
              "wang_p4_kd1", "wang_p4_kd2",
              "wang_nc_kd1", "wang_nc_kd2",
              "wang_p2_oe1", "wang_p2_oe2",
              "wang_nc_oe1", "wang_nc_oe2")

colnames(Wang_RPKM) <- c("id", "name", tmp_Name)

#reorder columns
Wang_RPKM <- Wang_RPKM[, c("id", "name",
                           "wang_nc_kd1", "wang_nc_kd2",
                           "wang_p2_kd1", "wang_p2_kd2", 
                           "wang_p4_kd1", "wang_p4_kd2",
                           "wang_p24_dkd1", "wang_p24_dkd2",
                           "wang_nc_oe1", "wang_nc_oe2",
                           "wang_p2_oe1", "wang_p2_oe2")]
```

##### Our_RPKM
```{r input our RPKM}
RPKM_files <- list.files("../input/polyA_RNAseq/", 
                         pattern = "*.gene_abund.txt")

#input data
RPKM <- inputStringTieRPKMfiles(RPKM_files,
                                RPKMDataPath = "../input/polyA_RNAseq/")

#remove ERCC annotations
RPKM <- RPKM[-grep("ERCC", RPKM$id),]

#use simplified sample names 
sampleName <- c("p2_MKO_1", "p2_MKO_7", "p2_MKO_9",
                "p2_MZKO_2", "p2_MZKO_4", "p2_MZKO_5", 
                "CTR1C_2", "CTR1C_3", "CTR2C_1", "CTR2C_2",
                "p4_MKO_1", "p4_MKO_2", "p4_MKO_5", 
                "p4_MZKO_10", "p4_MZKO_11", "p4_MZKO_6")

colnames(RPKM) <- c("id", "name", sampleName)

#reorder columns
RPKM <- RPKM[, c("id", "name",
                 "CTR1C_2", "CTR1C_3", "CTR2C_1", "CTR2C_2",
                 "p2_MKO_1", "p2_MKO_7", "p2_MKO_9",
                 "p2_MZKO_2", "p2_MZKO_4", "p2_MZKO_5",
                 "p4_MKO_1", "p4_MKO_2", "p4_MKO_5",
                 "p4_MZKO_6", "p4_MZKO_10", "p4_MZKO_11")]
```

Next obtain raw counts of repeats using TEtranscripts and use DESeq2 to normalize counts 
##### Organize counts
```{r input raw counts}
#input raw read counts
ReikCounts <- read.table(
  "../input/public/Reik_TEtranscript_out.cntTable",
  header = T, row.names = 1)
TronoCounts <- read.table(
  "../input/public/Trono_TEtranscripts_out.cntTable", 
  header = T, row.names = 1)
WangCounts <- read.table(
  "../input/public/Wang_TEtranscripts_out.cntTable",
  header = T, row.names = 1)
CTRcounts <- read.table(
  "../input/polyA_RNAseq/071721_TEtranscripts_out.cntTable",
  header = T, row.names = 1)
Dppa2Counts <- read.table(
  "../input/polyA_RNAseq/070721_TEtranscripts_out.cntTable", 
  header = T, row.names = 1)
Dppa4Counts <- read.table(
  "../input/polyA_RNAseq/Dppa4_single2C_TEtranscripts_out_062821.cntTable",
  header = T, row.names = 1)

#simplify column names

#05 = p2ko1, 12 = p2ko2, 23 = p4ko1, 29 = p4ko2,
#43 = dko, 57 = wt1, 58 = wt2
colnames(ReikCounts) <- c("reik_p2_ko1", "reik_p4_ko1", 
                         "reik_dko", "reik_p2_ko2",
                         "reik_p4_ko2", "reik_wt2", "reik_wt1")

#1A4=p2ko1, 1B1=p2ko2, 1B4=p2ko3
#1D7=p4ko1, 2D7=p4ko2, 3A5=p4ko3
#1B3=wt1, 1C3=wt2, 3B1=wt3
colnames(TronoCounts) <- c("trono_p2_ko3", "trono_p4_ko2", "trono_p2_ko1",
                           "trono_p4_ko1", "trono_p2_ko2", "trono_p4_ko3",
                           "trono_wt1", "trono_wt3", "trono_wt2")

colnames(WangCounts) <- c("wang_p2_oe1", "wang_p24_dkd1", "wang_p2_oe2",
                           "wang_p2_kd1", "wang_p4_kd2", "wang_p4_kd1",
                          "wang_p24_dkd2", "wang_p2_kd2", "wang_nc_oe1",
                          "wang_nc_kd2", "wang_nc_kd1", "wang_nc_oe2")

colnames(CTRcounts) <- c("CTR2C_2", "CTR2C_1", 
                         "CTR1C_3", "CTR1C_2")

colnames(Dppa2Counts) <- c("p2_MZKO_5", "p2_MZKO_4", "p2_MZKO_2",
                           "p2_MKO_1", "p2_MKO_9", "p2_MKO_7")
Dppa2Counts <- Dppa2Counts[, c("p2_MKO_1", "p2_MKO_7", "p2_MKO_9",
                               "p2_MZKO_2", "p2_MZKO_4", "p2_MZKO_5")]

colnames(Dppa4Counts) <- c("p4_MKO_5", "p4_MKO_2", "p4_MKO_1",
                           "p4_MZKO_10", "p4_MZKO_11", "p4_MZKO_6")
Dppa4Counts <- Dppa4Counts[, c("p4_MKO_1", "p4_MKO_2", "p4_MKO_5",
                               "p4_MZKO_6", "p4_MZKO_10", "p4_MZKO_11")]

#combine our Counts 
our_Counts <- cbind(CTRcounts, Dppa2Counts, Dppa4Counts)
```

##### Normalize counts
```{r DESeq2}
suppressMessages(library(DESeq2))

#Reik_lab normalized Counts
suppressMessages(
  Reik_normCounts <- countsToDEseq2FDR(
     counts = ReikCounts[, c("reik_wt1", "reik_wt2",
                            "reik_p2_ko1", "reik_p2_ko2", 
                            "reik_p4_ko1", "reik_p4_ko2",
                            "reik_dko")], 
     CGroup = 2, TGroup = 5)
 )

#Trono_lab normalized Counts
suppressMessages(
  Trono_normCounts <- countsToDEseq2FDR(
    counts = TronoCounts[, c("trono_wt1", "trono_wt3", "trono_wt2",
                             "trono_p2_ko1", "trono_p2_ko2", "trono_p2_ko3",
                             "trono_p4_ko1", "trono_p4_ko2", "trono_p4_ko3")], 
    CGroup = 3, TGroup = 6)
)

#Wang_lab normalized Counts
suppressMessages(
  Wang_normCounts <- countsToDEseq2FDR(
    counts = WangCounts[, c("wang_nc_kd1", "wang_nc_kd2",
                            "wang_p2_kd1", "wang_p2_kd2",
                            "wang_p4_kd1", "wang_p4_kd2", 
                            "wang_p24_dkd1", "wang_p24_dkd2",
                            "wang_nc_oe1",  "wang_nc_oe2",
                            "wang_p2_oe1",  "wang_p2_oe2"
                            )], 
    CGroup = 2, TGroup = 10)
)

#Our_lab normalized Counts
suppressMessages(
  our_normCounts <- countsToDEseq2FDR(
    counts = our_Counts[, c("CTR1C_2", "CTR1C_3",
                           "CTR2C_1", "CTR2C_2",
                           "p2_MKO_1", "p2_MKO_7", "p2_MKO_9",
                           "p2_MZKO_2", "p2_MZKO_4", "p2_MZKO_5",
                           "p4_MKO_1", "p4_MKO_2", "p4_MKO_5",
                           "p4_MZKO_6", "p4_MZKO_10", "p4_MZKO_11")], 
    CGroup = 4, TGroup = 12)
)
```

##### Combine gene/repeat

RPKM for genes and normalized counts (DESeq2) repeats
```{r combin gene repeats}

#Reik_lab
Reik_repeats <- Reik_normCounts[grep(":", Reik_normCounts$id),]
colnames(Reik_repeats)[14] <- "name"

Reik <- rbind(
  Reik_RPKM[, c("name", "reik_wt1", "reik_wt2", 
                "reik_p2_ko1", "reik_p2_ko2",
                "reik_p4_ko1", "reik_p4_ko2", "reik_dko")],
  Reik_repeats[, c("name", "reik_wt1", "reik_wt2", 
                "reik_p2_ko1", "reik_p2_ko2",
                "reik_p4_ko1", "reik_p4_ko2", "reik_dko")]
)

#Trono lab
Trono_repeats <- Trono_normCounts[grep(":", Trono_normCounts$id),]
colnames(Trono_repeats)[16] <- "name"

Trono <- rbind(
  Trono_RPKM[, c("name", "trono_wt1", "trono_wt2", "trono_wt3",
                 "trono_p2_ko1", "trono_p2_ko2", "trono_p2_ko3",
                 "trono_p4_ko1", "trono_p4_ko2", "trono_p4_ko3")],
  Trono_repeats[, c("name", "trono_wt1", "trono_wt2", "trono_wt3",
                 "trono_p2_ko1", "trono_p2_ko2", "trono_p2_ko3",
                 "trono_p4_ko1", "trono_p4_ko2", "trono_p4_ko3")]
)

#Wang lab
Wang_repeats <- Wang_normCounts[grep(":", Wang_normCounts$id),]
colnames(Wang_repeats)[19] <- "name"

Wang <- rbind(
  Wang_RPKM[, c("name", "wang_nc_kd1", "wang_nc_kd2",
                "wang_p2_kd1", "wang_p2_kd2",  
                "wang_p4_kd1", "wang_p4_kd2",
                "wang_p24_dkd1", "wang_p24_dkd2",
                "wang_nc_oe1", "wang_nc_oe2",  
                "wang_p2_oe1", "wang_p2_oe2")],
  Wang_repeats[, c("name", "wang_nc_kd1", "wang_nc_kd2",
                "wang_p2_kd1", "wang_p2_kd2",  
                "wang_p4_kd1", "wang_p4_kd2",
                "wang_p24_dkd1", "wang_p24_dkd2",
                "wang_nc_oe1", "wang_nc_oe2",  
                "wang_p2_oe1", "wang_p2_oe2")]
)

#Zhang lab
Our_repeats <- our_normCounts[grep(":", our_normCounts$id),]
colnames(Our_repeats)[23] <- "name"

Our <- rbind(
  RPKM[, c("name", "CTR1C_2", "CTR1C_3", "CTR2C_1", "CTR2C_2",   
           "p2_MKO_1", "p2_MKO_7", "p2_MKO_9", "p2_MZKO_2",
           "p2_MZKO_4", "p2_MZKO_5", "p4_MKO_1", "p4_MKO_2",
           "p4_MKO_5", "p4_MZKO_6",  "p4_MZKO_10", "p4_MZKO_11")],
  Our_repeats[, c("name", "CTR1C_2", "CTR1C_3", "CTR2C_1", "CTR2C_2",   
           "p2_MKO_1", "p2_MKO_7", "p2_MKO_9", "p2_MZKO_2",
           "p2_MZKO_4", "p2_MZKO_5", "p4_MKO_1", "p4_MKO_2",
           "p4_MKO_5", "p4_MZKO_6",  "p4_MZKO_10", "p4_MZKO_11")]
)
```

##### heatmap
```{r heatmap}
gene2C <- c("Zscan4c", "Zscan4d", 
           "Zfp352", 
           "Tcstv3",
           "Pramef25",
           "Kdm4dl",
           "Tdpoz3",
           "Tdpoz4",
           "MERVL-int:ERVL:LTR", "MT2_Mm:ERVL:LTR", "GSAT_MM:Satellite:Satellite"
           )

suppressMessages(library(pheatmap))
suppressMessages(library(RColorBrewer))
suppressMessages(library(ComplexHeatmap))
suppressMessages(library(circlize))

#Reik_lab
Reik2C <- Reik[which(Reik$name %in% gene2C),]
Reik2Cmat <- log2(as.matrix(Reik2C[, c(2:8)]+0.1))
row.names(Reik2Cmat) <- Reik2C$name

#Trono_lab
Trono2C <- Trono[which(Trono$name %in% gene2C),]
Trono2Cmat <- log2(as.matrix(Trono2C[, c(2:10)]+0.1))
row.names(Trono2Cmat) <- Trono2C$name

#Wang_lab
Wang2C <- Wang[which(Wang$name %in% gene2C),]
Wang2Cmat1 <- log2(as.matrix(Wang2C[, c(2:9)]+0.1))
Wang2Cmat2 <- log2(as.matrix(Wang2C[, c(10:13)]+0.1))
row.names(Wang2Cmat1) <- Wang2C$name
row.names(Wang2Cmat2) <- Wang2C$name

#our lab 
Our2C <- Our[which(Our$name %in% gene2C),]
Our2Cmat <- log2(as.matrix(Our2C[, c(2:17)]+0.1))
row.names(Our2Cmat) <- Our2C$name

#Heatmap
pdf(file = paste0("../figure/", format(Sys.Date(), "%y%m%d"), 
                   "_2CgenesInESCsEmbryos.pdf"), width = 6, height = 6)
breaksList = seq(-2, 2, by = 0.1)

pheatmap(Reik2Cmat, scale = "row",
         cluster_cols = F, cluster_rows = F, treeheight_row = 0,
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(length(breaksList)),
         breaks = breaksList)

pheatmap(Trono2Cmat, scale = "row",
         cluster_cols = F, cluster_rows = F, treeheight_row = 0,
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(length(breaksList)),
         breaks = breaksList)

pheatmap(Wang2Cmat1, scale = "row",
         cluster_cols = F, cluster_rows = F, treeheight_row = 0,
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(length(breaksList)),
         breaks = breaksList)

pheatmap(Wang2Cmat2, scale = "row",
         cluster_cols = F, cluster_rows = F, treeheight_row = 0,
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(length(breaksList)),
         breaks = breaksList)

pheatmap(Our2Cmat, scale = "row",
         cluster_cols = F, cluster_rows = F, treeheight_row = 0,
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(length(breaksList)),
         breaks = breaksList)
dev.off()
```




